---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tmutilsr

tmutilsr provides text processing utilities and extends basic R text mining packages,
such as topicmodels and stm, with commonly used processing functions

**Topic modeling utilities (basic package topicmodels)**

* `lda_trim_beta` - trims LDA beta table by terms and/or topics
* `lda_table_topics` - spreads the term distribution over topics by topic
* `lda_list_docs` - lists documents by their lead topic

**Text cleanup utilities**

* `tm_stem2readable` - transforms stemmed words to a readable format
* `tm_make_corpus` - Create a clean, stemmed and tidy corpus

**URL scraping utilities**

* `scrape_text` - scrapes text of a specific node tag in a given URL

## Installation

The package is currently not availabel on CRAN

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
## basic example code
```

